---
title: "Life Expectancy Analysis"
author: "Lijju Mathew, Blake Freeman, Aniketh Vankina"
date: "June 4, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

install.packages("model4you")

```{r library}
library(dvmisc)
library(rgl)
library(Metrics)
library(tidyverse)
library(ggplot2)
library(caTools)
library(dplyr)
library(magrittr)
library(readr)
library(survival)
library(nlme)
library(gridExtra) #grid.arrange()
library(class)
library(forcats)
library(MASS)
library(GGally)
library(tidyr)
library(maps)
library(mapproj)
library(stringr)
library(rmarkdown)
library(knitr)
library(jsonlite)
library(RCurl)
library(class)
library(httr)
library(mice)
library(corrplot)
library(GoodmanKruskal) # GKtauDataframe()
library(glmnet)
library(car)
library(moderndive) # get_regression_table()
library(faraway)
library(randomForest)
library(e1071)
library(naivebayes)
library(caret)
library(SuperLearner)
library(psych)
library(tm)
library(VIM)
library(model4you)


```

============
OBJECTIVE 1
============

```{r datacleanup}
who_data <- read.csv(file.choose())
#who_data <- read.csv("/Users/lijjumathew/Library/Mobile Documents/com~apple~CloudDocs/Lijju/SMU/Courses/Applied Statistics/Project/Life Expectancy Data.csv")
who_data <- who_data %>% filter(Year == 2014)
head(who_data)
summary(who_data)
str(who_data)

# Find the no of attributes with missing values
sort(sapply(who_data, function(x) sum(is.na(x))), decreasing = T)

#Missing data and percentage plot
missing.values <- who_data %>%
  gather(key = "key", value = "val") %>%
  dplyr::mutate(isna = is.na(val)) %>%
  dplyr::group_by(key) %>%
  dplyr::mutate(total = n()) %>%
  dplyr::group_by(key, total, isna)%>%
  dplyr::summarise(num.isna = n())%>%
  dplyr::mutate(pct = num.isna / total * 100)

levels <- (missing.values %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
      ggplot() + geom_bar(aes(x = reorder(key, desc(pct)), y = pct, fill=isna), stat = 'identity', alpha=0.8) +
      scale_x_discrete(limits = levels) +
      scale_fill_manual(name = "", values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
      coord_flip() + 
      labs(title = "Percentage of missing values", x = 'Variable', y = "% of missing values")
percentage.plot

tempData <- mice(who_data,m=1,maxit=0,meth='fastpmm',seed=500)
who_imp <- complete(tempData,1)

sort(sapply(who_imp, function(x) sum(is.na(x))), decreasing = T)
glimpse(who_imp)
```

```{r split}
set.seed(100)
split_percent = .70
trainIndices = sample(1:dim(who_imp)[1],round(split_percent * dim(who_imp)[1]))
train = who_imp[trainIndices,]
test = who_imp[-trainIndices,]
summary(train)
```


dev.off()
=======
```{r}
#For Git
#EDA Aniketh f
library(ISLR)
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(tidyr)
library(magrittr)
library(plyr)

life = read.csv(file.choose(), header = TRUE)
summary(life)
dplyr::glimpse(life)
attach(life)
#Life expectancy based on developing and developed countries
t(aggregate(Life.expectancy ~ Status, data = life, summary))


#Does various predicting factors which has been chosen initially really affect the Life expectancy? What are the predicting variables actually affecting the life expectancy?
library(car)
who_imp = who_imp[,-3]
who_imp = who_imp[,-1]
full.model = lm(Life.expectancy~., data = who_imp)
vif(full.model)[,3]^2

alias(lm(Life.expectancy~., data = who_imp))

#Should a country having a lower life expectancy value(<65) increase its healthcare expenditure in order to improve its average lifespan?
life %>% select(Country, Life.expectancy, Total.expenditure, percentage.expenditure) %>% filter(Life.expectancy < 65) 

#How does Infant and Adult mortality rates affect life expectancy?
par(mfrow = c(1,2))
plot(Adult.Mortality, Life.expectancy, xlab = "Adult Mortality", ylab = "Life Expectancy")
new = data.frame(Adult.Mortality = seq(30, 600,1))
lines(seq(30,600,1), predict(lm(Life.expectancy~Adult.Mortality), newdata = new), col = "red", lwd = 4)

plot(infant.deaths, Life.expectancy)
new2 = data.frame(infant.deaths = seq(30,600,1))
lines(seq(30,600,1), predict(lm(Life.expectancy~infant.deaths), newdata = new2), col = "red", lwd = 4)
#Does Life Expectancy has positive or negative correlation with eating habits, lifestyle, exercise, smoking, drinking alcohol etc.

#What is the impact of schooling on the lifespan of humans?
plot(Schooling, Life.expectancy, xlab = "Schooling", ylab = "Life Expectancy")
new3 = data_frame(Schooling = seq(1, 30,1))
lines(seq(1,30,1), predict(lm(Life.expectancy~Schooling), newdata = new3), col = "red", lwd = 4) 
#There is a postive corelation between life expectancy and schooling

#Does Life Expectancy have positive or negative relationship with drinking alcohol?
plot(Alcohol, Life.expectancy, xlab = "Alcohol", ylab = "Life Expectancy")
new3 = data_frame(Alcohol = seq(0.1, 30,.1))
lines(seq(0.1,30,.1), predict(lm(Life.expectancy~Alcohol), newdata = new3), col = "red", lwd = 4) 
#Do densely populated countries tend to have lower life expectancy?

#What is the impact of Immunization coverage on life Expectancy?


```


```{r EDA EDA EDA}
summary(who_data)
who_imp_conti <- who_imp[, !sapply(who_imp, is.factor)]
who_imp_categ <- who_imp[, sapply(who_imp, is.factor)]

# Box plots to find outliers
boxplot(who_imp_conti[,2:20])
boxplot(who_imp_conti$Population)
who_imp[which.max(who_imp$Population),]

# ScatterPlot
head(who_imp_conti)
pairs(who_imp_conti[,2:20], pch=19)

# Computing the p value of correlations
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
p.mat <- cor.mtest(who_imp_conti[,2:20])
correlation <- cor(who_imp_conti[,2:20])
# Correlation plot with significance level of 0.05
corrplot(correlation, type="upper", order="hclust", p.mat = p.mat, sig.level = 0.05)

cat_cor<- GKtauDataframe(who_imp_categ)
plot(cat_cor, corrColors = "blue")
```

EDA
1. Out of total 183 countries, 151 countries are developing  and 32 are developed
2. Outliers - Population of India seemed as an outlier, but that is a fact.
3. Correlation 
    Percentage.expenditure - GDP
    Polio - Hepatitis.B, Diptheria
    Life.expectancy - Income.composition.of.resources, Schooling, Adult.Mortality, HIV.AIDS 
    Income.composition.of.resources - Schooling, Adult.Mortality, HIV.AIDS 
    Schooling - Adult.Mortality
    thinness..1.19.years - thinness..5.9.years
    Population - infant.deaths, under.five.deaths
    infant.deaths - under.five.deaths

dev.off()

```{r Full Model with all predictors}
par(mfrow=c(2,2))
model_full = lm(Life.expectancy ~ Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + Hepatitis.B + 
                Measles + BMI + under.five.deaths + Polio + Total.expenditure + Diphtheria + HIV.AIDS + GDP + 
                Population + thinness..1.19.years + thinness.5.9.years + Income.composition.of.resources + 
                Schooling,data=train)
scope_formula = formula(model_full)
scope_formula
model_full_selection = step(object=model_full, scope=scope_formula, direction="backward")
summary(model_full_selection)
plot(model_full_selection)

model_full_selection_aic = AIC(model_full_selection)
model_full_selection_bic = BIC(model_full_selection)
model_full_selection_aic # AIC of the model using forward selection method
model_full_selection_bic
ASE=rss(model_full_selection)/nrow(train)
ASE
```
==================================================================================
From the Full Model the below are the significant predictors - Forward Selection
==================================================================================

Adult.Mortality                 -1.730e-02  4.586e-03  -3.772 0.000263 ***
infant.deaths                    3.647e-01  1.102e-01   3.309 0.001270 ** 
Alcohol                          1.928e-01  9.368e-02   2.058 0.041997 *  
under.five.deaths               -2.948e-01  8.244e-02  -3.576 0.000521 ***
HIV.AIDS                        -5.697e-01  2.959e-01  -1.926 0.056758 .  
Income.composition.of.resources  3.141e+01  5.572e+00   5.638 1.37e-07 ***

Residual standard error: 3.323 on 109 degrees of freedom
Multiple R-squared:  0.8817,	Adjusted R-squared:  0.8622 
AIC=690.1365  BIC=747.1771
==================================================================================
From the Full Model the below are the significant predictors - Backward Selection
==================================================================================

Adult.Mortality                 -0.019037   0.004314  -4.413 2.24e-05 ***
infant.deaths                    0.346787   0.091338   3.797 0.000231 ***
Alcohol                          0.190732   0.084952   2.245 0.026588 *  
under.five.deaths               -0.278161   0.072770  -3.822 0.000211 ***
Total.expenditure                0.208568   0.106121   1.965 0.051680 .  
HIV.AIDS                        -0.542359   0.271797  -1.995 0.048258 *  
Income.composition.of.resources 32.084721   3.092474  10.375  < 2e-16 ***

Residual standard error: 3.238 on 120 degrees of freedom
Multiple R-squared:  0.8764,	Adjusted R-squared:  0.8692  
AIC=673.8125  BIC=699.4808
==================================================================================
From the Full Model the below are the significant predictors - Stepwise Selection
==================================================================================

Adult.Mortality                 -0.019037   0.004314  -4.413 2.24e-05 ***
infant.deaths                    0.346787   0.091338   3.797 0.000231 ***
Alcohol                          0.190732   0.084952   2.245 0.026588 *  
under.five.deaths               -0.278161   0.072770  -3.822 0.000211 ***
Total.expenditure                0.208568   0.106121   1.965 0.051680 .  
HIV.AIDS                        -0.542359   0.271797  -1.995 0.048258 *  
Income.composition.of.resources 32.084721   3.092474  10.375  < 2e-16 ***

Residual standard error: 3.238 on 120 degrees of freedom
Multiple R-squared:  0.8764,	Adjusted R-squared:  0.8692
AIC=673.8125  BIC=699.4808
==================================================================================

dev.off()
```{r reduced model1- from manual selection, echo=FALSE}
par(mfrow=c(2,2))
reduced_model_1 <- lm(Life.expectancy~Schooling+Income.composition.of.resources+Alcohol+ HIV.AIDS, data=train)
get_regression_table(reduced_model_1)
summary(reduced_model_1)
# Model validation using test set
test$Life.expectancy.Predicted <- predict(reduced_model_1, subset(test,select=-c(Country,Life.expectancy)))
test.correlation <- round(cor(test$Life.expectancy, test$Life.expectancy.Predicted),4)
test.RMSE <- round(sqrt(mean(test$Life.expectancy-test$Life.expectancy.Predicted)^2),4)
c(correlation = test.correlation, RMSE = test.RMSE)
vif(reduced_model_1)

reduced_model_1_aic = AIC(reduced_model_1)
reduced_model_1_bic = BIC(reduced_model_1)
reduced_model_1_aic # AIC of the model using forward selection method
reduced_model_1_bic

plot(reduced_model_1)
# Scatterplot
pairs(data.frame(train$Life.expectancy,train$Schooling,train$Alcohol,train$Income.composition.of.resources,train$HIV.AIDS), pch=19)
ASE=rss(reduced_model_1)/nrow(train)
ASE
```

From the Reduced Model (manual selection) the below are the significant predictors 
1.Income.composition.of.resources
2.HIV.AIDS
Residual standard error: 3.708 on 123 degrees of freedom
Multiple R-squared:  0.8159,	Adjusted R-squared:   0.81 
correlation        RMSE 
     0.8378      0.5866 
VIF
Schooling = 6.275441                 
Income.composition.of.resources = 6.636833                      
Alcohol = 1.571182         
HIV.AIDS = 1.254619
AIC=705.623  BIC=722.7352

dev.off()

```{r reduced model2 - from stepwise selection, echo=FALSE}
par(mfrow=c(2,2))
reduced_model_2 <- lm(Life.expectancy~Adult.Mortality+infant.deaths+Alcohol+under.five.deaths+Total.expenditure+ HIV.AIDS+Income.composition.of.resources, data=train)
get_regression_table(reduced_model_2)
summary(reduced_model_2)
# Model validation using test set
test$Life.expectancy.Predicted <- predict(reduced_model_2, subset(test,select=-c(Country,Life.expectancy)))
test.correlation <- round(cor(test$Life.expectancy, test$Life.expectancy.Predicted),4)
test.RMSE <- round(sqrt(mean(test$Life.expectancy-test$Life.expectancy.Predicted)^2),4)
c(correlation = test.correlation, RMSE = test.RMSE)
vif(reduced_model_2)
plot(reduced_model_2)
reduced_model_2_aic = AIC(reduced_model_2)
reduced_model_2_bic = BIC(reduced_model_2)
reduced_model_2_aic # AIC of the model using forward selection method
reduced_model_2_bic
# Scatterplot
pairs(data.frame(train$Life.expectancy, train$Adult.Mortality,train$Total.expenditure,train$Income.composition.of.resources,train$HIV.AIDS), pch=19)
ASE=rss(reduced_model_2)/nrow(train)
ASE
```
==================================================================================
From the Reduced Model (stepwise selection) the below are the significant predictors 
==================================================================================

Adult.Mortality                 -0.019037   0.004314  -4.413 2.24e-05 ***
infant.deaths                    0.346787   0.091338   3.797 0.000231 ***
Alcohol                          0.190732   0.084952   2.245 0.026588 *  
under.five.deaths               -0.278161   0.072770  -3.822 0.000211 ***
Total.expenditure                0.208568   0.106121   1.965 0.051680 .  
HIV.AIDS                        -0.542359   0.271797  -1.995 0.048258 *  
Income.composition.of.resources 32.084721   3.092474  10.375  < 2e-16 ***

Residual standard error: 3.238 on 120 degrees of freedom
Multiple R-squared:  0.8764,	Adjusted R-squared:  0.8692 
correlation        RMSE 
     0.7630      0.4363 
VIF
Adult.Mortality - 2.823806 
infant.deaths - 800.207011
Alcohol - 1.497704
under.five.deaths - 805.376827
Total.expenditure - 1.116503 
HIV.AIDS - 1.811741
Income.composition.of.resources - 2.760631
AIC=673.8125  BIC=699.4808
==================================================================================

dev.off()

```{r LASSO,  echo=FALSE}
set.seed(100)
#Formatting data for GLM net
x=model.matrix(Life.expectancy ~ Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + Hepatitis.B + 
                Measles + BMI + under.five.deaths + Polio + Total.expenditure + Diphtheria + HIV.AIDS + GDP + 
                Population + thinness..1.19.years + thinness.5.9.years + Income.composition.of.resources + 
                Schooling, train)[,-1]
Life.expectancy~Schooling+Income.composition.of.resources+Total.expenditure+percentage.expenditure+BMI+Alcohol
y=train$Life.expectancy

xtest<-model.matrix(Life.expectancy ~ Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + Hepatitis.B + 
                Measles + BMI + under.five.deaths + Polio + Total.expenditure + Diphtheria + HIV.AIDS + GDP + 
                Population + thinness..1.19.years + thinness.5.9.years + Income.composition.of.resources + 
                Schooling,test)[,-1]
ytest<-test$Life.expectancy

grid=10^seq(10,-2, length =100)
lasso.mod=glmnet(x,y,alpha=1, lambda =grid)

cv.out=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
plot(cv.out)
bestlambda<-cv.out$lambda.min  #Optimal penalty parameter.  You can make this call visually.
lasso.pred=predict (lasso.mod ,s=bestlambda ,newx=xtest)

testMSE_LASSO<-mean((ytest-lasso.pred)^2)
testMSE_LASSO
coef(lasso.mod,s=bestlambda)
```
================
Lasso Predictors 
================
Adult.Mortality                 -0.0191246691
Alcohol                          0.1290622227
Total.expenditure                0.0559652197
Diphtheria                       0.0002095938
HIV.AIDS                        -0.4152214620
thinness..1.19.years            -0.0494267615
Income.composition.of.resources 34.2057499766


dev.off()

```{r reduced model3 - from Lasso selection, echo=FALSE}
par(mfrow=c(2,2))
reduced_model_3 <- lm(Life.expectancy ~ Adult.Mortality + Alcohol + Total.expenditure + Diphtheria + HIV.AIDS + thinness..1.19.years + Income.composition.of.resources,data=train)
get_regression_table(reduced_model_3)
summary(reduced_model_3)
# Model validation using test set
test$Life.expectancy.Predicted <- predict(reduced_model_3, subset(test,select=-c(Country,Life.expectancy)))
test.correlation <- round(cor(test$Life.expectancy, test$Life.expectancy.Predicted),4)
test.RMSE <- round(sqrt(mean(test$Life.expectancy-test$Life.expectancy.Predicted)^2),4)
c(correlation = test.correlation, RMSE = test.RMSE)
vif(reduced_model_3)
plot(reduced_model_3)
reduced_model_3_aic = AIC(reduced_model_3)
reduced_model_3_bic = BIC(reduced_model_3)
reduced_model_3_aic # AIC of the model using forward selection method
reduced_model_3_bic
summary(reduced_model_3)
ASE=rss(reduced_model_3)/nrow(train)
ASE
```

dev.off()


====================================================================================
From the Reduced Model (lasso selection) the below are the significant predictors 
====================================================================================

Adult.Mortality                 -0.019658   0.004567  -4.305 3.43e-05 ***
Alcohol                          0.190417   0.089385   2.130   0.0352 *  
Total.expenditure                0.158346   0.114060   1.388   0.1676    
Diphtheria                       0.014960   0.014128   1.059   0.2918    
HIV.AIDS                        -0.631963   0.284766  -2.219   0.0284 *  
thinness..1.19.years            -0.112665   0.096625  -1.166   0.2459    
Income.composition.of.resources 33.338972   3.397006   9.814  < 2e-16 ***

Residual standard error: 3.402 on 120 degrees of freedom
Multiple R-squared:  0.8636,	Adjusted R-squared:  0.8556 
correlation        RMSE 
     0.8026      0.2577
VIF
Adult.Mortality - 2.867694 
Alcohol - 1.502603
Total.expenditure - 1.168843 
Diphtheria - 1.149265
HIV.AIDS - 1.802256
thinness..1.19.years - 1.544284
Income.composition.of.resources - 3.018706
AIC=686.4176  BIC=712.0859
==================================================================================

```{r model4 - Life Expectancy differences between developed and devloping countries, echo=FALSE}
par(mfrow=c(2,2))
reduced_model_4 <- lm(Life.expectancy ~ Status ,data=who_data)
get_regression_table(reduced_model_4)
summary(reduced_model_4)
```
====================================================================================
From the Linear model Life Expectancy for Developing vs Developed
====================================================================================
(Intercept)        81.138      1.298  62.493  < 2e-16 ***
StatusDeveloping  -11.636      1.429  -8.141 6.17e-14 ***

1. The P value of dummy variable StatusDeveloping is very significant suggesting that 
  there is statistical evidence of difference in average life expectancy between developing and developed countries.
2. The average life expectancy in developed countries is estimated to be 81.138
3. The average life expectancy in developing countries is 11.638 years less than developed countries.


dev.off()

```{r Final Model1 , echo=FALSE}
set.seed(100)
par(mfrow=c(2,2))
final_model1 <- lm(Life.expectancy ~ Adult.Mortality +  Alcohol + Total.expenditure + HIV.AIDS + Income.composition.of.resources + thinness..1.19.years,data=train)
get_regression_table(final_model1)
summary(final_model1)
# Model validation using test set
test$Life.expectancy.Predicted <- predict(final_model1, subset(test,select=-c(Life.expectancy)))
test.correlation <- round(cor(test$Life.expectancy, test$Life.expectancy.Predicted),4)
test.RMSE <- round(sqrt(mean(test$Life.expectancy-test$Life.expectancy.Predicted)^2),4)
c(correlation = test.correlation, RMSE = test.RMSE)
vif(final_model1)
plot(final_model1)
final_model1_aic = AIC(final_model1)
final_model1_bic = BIC(final_model1)
final_model1_aic # AIC of the model using forward selection method
final_model1_bic
summary(final_model1)
ASE=rss(final_model1)/nrow(train)
ASE

```



```{r Final Model2 , echo=FALSE}
par(mfrow=c(2,2))
final_model2 <- lm(Life.expectancy ~ Adult.Mortality +  Total.expenditure + Income.composition.of.resources ,data=train)
get_regression_table(final_model2)
summary(final_model2)
# Model validation using test set
test$Life.expectancy.Predicted <- predict(final_model2, subset(test,select=-c(Life.expectancy)), interval="confidence")
test.correlation <- round(cor(test$Life.expectancy, test$Life.expectancy.Predicted),4)
test.RMSE <- round(sqrt(mean(test$Life.expectancy-test$Life.expectancy.Predicted)^2),4)
c(correlation = test.correlation, RMSE = test.RMSE)
vif(final_model2)
plot(final_model2)
final_model2_aic = AIC(final_model2)
final_model2_bic = BIC(final_model2)
final_model2_aic # AIC of the model using forward selection method
final_model2_bic
summary(final_model2)
ASE=rss(final_model2)/nrow(train)
ASE
```


dev.off()

===========
OBJECTIVE 2
===========

```{r knn with all the predictors}
# Train model with knn and get importance of predictors
control <- trainControl(method="repeatedcv", number=10, repeats=3)
set.seed(100)
model.knn <- train( Life.expectancy ~ ., data=who_imp[,-c(2)], method="knn", trControl=control)
#Top 10 predictor ranking
importance.knn <- varImp(model.knn, scale=FALSE)
rank.knn <- importance.knn$importance
write.csv(rank.knn, "rank.knn.csv")
rank.knn <- read.csv("rank.knn.csv", header=TRUE)
colnames(rank.knn) <- c("Predictors", "Importance")
rank.knn <- rank.knn[order(rank.knn$Importance, decreasing = TRUE),]
ggplot(rank.knn[1:20,], aes(x=reorder(Predictors, Importance),y=Importance)) + geom_bar(stat = "identity") + coord_flip() + labs(title="Importance of Predictors", x="Predictors", y="Importance") +theme(axis.text.x=element_text(hjust=0.5, vjust=0.5, size = 12))+theme(axis.text.y=element_text(size = 12))
``` 


```{r knn with all the predictors}
# Train model with knn and get importance of predictors
control <- trainControl(method="repeatedcv", number=10, repeats=3)
set.seed(100)
model.knn <- train(Life.expectancy ~ Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + Hepatitis.B + 
                Measles + BMI + under.five.deaths + Polio + Total.expenditure + Diphtheria + HIV.AIDS + GDP + 
                Population + thinness..1.19.years + thinness.5.9.years + Income.composition.of.resources + 
                Schooling, data=who_imp[,-c(2)], method="knn", trControl=control)
#Top 10 predictor ranking
importance.knn <- varImp(model.knn, scale=FALSE)
rank.knn <- importance.knn$importance
write.csv(rank.knn, "rank.knn.csv")
rank.knn <- read.csv("rank.knn.csv", header=TRUE)
colnames(rank.knn) <- c("Predictors", "Importance")
rank.knn <- rank.knn[order(rank.knn$Importance, decreasing = TRUE),]
ggplot(rank.knn[1:20,], aes(x=reorder(Predictors, Importance),y=Importance)) + geom_bar(stat = "identity") + coord_flip() + labs(title="Importance of Predictors", x="Predictors", y="Importance") +theme(axis.text.x=element_text(hjust=0.5, vjust=0.5, size = 12))+theme(axis.text.y=element_text(size = 12))
``` 

Life.expectancy ~ Adult.Mortality +  Total.expenditure +  Income.composition.of.resources
Adult.Mortality +  Alcohol + Total.expenditure + HIV.AIDS + Income.composition.of.resources + thinness..1.19.years

```{r}
library(dplyr)
lifeReduced = who_imp %>% select(Life.expectancy, Adult.Mortality, Total.expenditure, Income.composition.of.resources)
lifeFull = who_imp %>% select(Life.expectancy, Adult.Mortality, Alcohol, Total.expenditure, HIV.AIDS, Income.composition.of.resources, thinness..1.19.years)
```

```{r knn life reduced}
set.seed(100)
iterations = 10
numks = 30
splitPerc = .70
masterAcc = matrix(nrow = iterations, ncol = numks)
for(j in 1:iterations)
{
   trainIndices = sample(1:dim(lifeReduced)[1],round(splitPerc * dim(lifeReduced)[1]))
    train = lifeReduced[trainIndices,]
    test = lifeReduced[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(2,3,4)],test[,c(2,3,4)],train$Life.expectancy, prob = TRUE, k = i)
    u <- union(classifications,test$Life.expectancy)
    t <- table(factor(classifications, u), factor(test$Life.expectancy, u))
    CM = confusionMatrix(t)
    masterAcc[j,i] = CM$overall[1]
  }
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
which.max(MeanAcc)
max(MeanAcc)
```


```{r knn life full}
set.seed(100)
iterations = 10
numks = 30
splitPerc = .70
masterAcc = matrix(nrow = iterations, ncol = numks)
for(j in 1:iterations)
{
   trainIndices = sample(1:dim(lifeFull)[1],round(splitPerc * dim(lifeFull)[1]))
    train = lifeFull[trainIndices,]
    test = lifeFull[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(2,3,4,5,6,7)],test[,c(2,3,4,5,6,7)],train$Life.expectancy, prob = TRUE, k = i)
    u <- union(classifications,test$Life.expectancy)
    t <- table(factor(classifications, u), factor(test$Life.expectancy, u))
    CM = confusionMatrix(t)
    masterAcc[j,i] = CM$overall[1]
  }
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
which.max(MeanAcc)
max(MeanAcc)
```



